Are we learning to build better models at the cost of better decisions?

### 2021-03-01 manuscript/appendix.Rmd
data cache

@@

The optimization problem can then be solved by dynamic programming, using the recursive Bellman equation as ,

@@

Bear in mind that in the example presented here, SDP merely serves the role of a convenient and established way for a manager to determine the optimal action, given some model estimation.  Whether we use SDP or some other way to determine the optimal policy, given the model, is immaterial to the conclusion.  The issue at hand is how we determine the model in the first place.

Given the function $f$ with known parameters, it is straight forward to determine the optimal harvest policy by stochastic dynamic programming (SDP, see @Mangel1985; @Marescot2013).

@@

policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp(P, reward, discount = discount, 
                epsilon = 0.01, max_iter = 2000, V0 = rep(0,dim(P)[[1]]))
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")
```

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
am %>%  ggplot(aes(time, states[state], col = belief)) + 
  geom_line() + 
  geom_point() +
  scale_colour_gradient(limits = c(0,1), low = pal[1], high = pal[4])
```

```{r include = FALSE}
as_tibble(am1$posterior) %>% 
  mutate(time = seq_along(V1)) %>%
  select(time = time, "1" = V1, "2" = V2) %>% 
  pivot_longer(c("1", "2"), values_to="belief", names_to = "model") %>%
  filter(time < 20) %>%
  ggplot(aes(time, belief, col=model)) + geom_line(col = pal[4])
```

@@

sims <- 
  map_dfr(names(transitions), 
          function(m){
            policy <- policies %>% filter(model == m) %>% pull(policy)
            map_dfr(1:reps, simulate_policy, policy = policy, .id = "reps")
          },
          .id = "model"
         )
```

@@

```{r}
predictions <- 
  stepahead_unfished %>% 
  select(time, model, reps, expected, low, high, true, score) %>% 
  mutate(scenario = "A_unfished") %>% 
  bind_rows(stepahead_fished  %>% 
          select(time, model, reps, expected, low, high, true, score) %>% 
          mutate(scenario = "B_fished")
  ) %>% 
  mutate(model = as.character(model))
```

# Plotting

Having computed a single data frame of simulation data under both optimally fished (according to each model) and unfished scenarios for each model, along with the proper scores for the step-ahead forecasts a la @Gneiting2007, we have all the results necessary to generate the figures presented in the main paper.

```{r fig1a}
fig1ab <- predictions %>%
  filter(reps == "2", time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, 
             alpha = 0.6) + 
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "A. Without fishing",
    B_fished = "B. Managed harvest"))) +
  ylab("stock size")

@@

Likewise we can plot the data on proper scores associated with each prediction of each model:

```{r}
fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C.",
    B_fished = "D."))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

## Forecast ecological and economic performance

The ecological performance of the model can easily be visualized by plotting
the results of each simulation.

```{r plot_sims}
fig2a <- 
  sims %>% 
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  ggtitle("A") + labs(subtitle = "Ecological")
```

@@

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

fig2b <- 
npv_df %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  ggtitle("B.") + labs(subtitle = "Economic")

@@

Plotting the model functions themselves are shown in Fig 3a:

```{r plot_models}
d <- 
  map_dfc(models, function(f) f(states) - states) %>%
  mutate(state = states)

fig3a <- 
  d %>% pivot_longer(names(models), "model") %>%
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")
```

We can easily plot the optimal policy derived from each model, as shown
in Fig3b.

```{r plot_policies}
fig3b <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")
```

>>>>

This appendix provides a fully reproducible walk-through of the implementation of the analyses considered in this paper.  None of the methods applied here: Stochastic Dynamic Programming, Iterative Forecasts assessed with Proper Scoring Rules, or Adaptive Management, are novel

This appendix includes computer code in the R language, which can also be found in the corresponding R-Markdown document in the paper's GitHub repository, <https://github.com/cboettig/bad-forecast-good-decision>.  While computer code can be more verbose and difficult to read than mathematical formulae presented here, it is also less ambiguous.

@@

The optimization problem can then be solved by dynamic programming, using the recursive Bellman equation as,

@@

Bear in mind that in the example presented here, SDP merely serves the role of a convenient and established way for a manager to determine the optimal action, given some model estimation.  Whether we use SDP or some other way to determine the optimal policy, given the model, is immaterial to the conclusion.

@@

policies <- 
  map_dfr(transitions, 
          function(P){
    soln <- mdp(P, reward, discount = discount, 
                epsilon = 0.01, max_iter = 2000, V0 = rep(0,dim(P)[[1]]))
    escapement <- states - actions[soln$policy]
    tibble(states, policy = soln$policy, escapement)
    }, 
  .id = "model")

write_csv(policies, "../data/policies.csv")
```

@@

sims <- 
  map_dfr(names(transitions), 
          function(m){
            policy <- policies %>% filter(model == m) %>% pull(policy)
            map_dfr(1:reps, simulate_policy, policy = policy, .id = "reps")
          },
          .id = "model"
         )

write_csv(sims, "../data/sims.csv")
```

@@

```{r}
predictions <- 
  stepahead_unfished %>% 
  select(time, model, reps, expected, low, high, true, score) %>% 
  mutate(scenario = "A_unfished") %>% 
  bind_rows(stepahead_fished  %>% 
          select(time, model, reps, expected, low, high, true, score) %>% 
          mutate(scenario = "B_fished")
  ) %>% 
  mutate(model = as.character(model))

write_csv(predictions, "../data/predictions.csv")
```

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
write_csv(am, "../data/am.csv")

@@

### Additional calculations for plots

@@

write_csv(npv_df, "../data/npv_df.csv")
```

```{r}
# tabular comparisons
npv <- npv_df %>% group_by(model) %>% summarize(npv = max(mean_npv))

am_npv <- sum(am$value * discount ^ am$time)
am_economics_percent <- round(am_npv / npv[[1,"npv"]] * 100)

mean_state <- sims %>% group_by(model) %>% summarize(state = mean(state))
am_ecology_percent <- round(mean(states[am$state])/mean_state[[1, "state"]]*100)
```

Under adaptive management, the manager realizes only `r am_economics_percent`%
of the economic value that would be achieved under Model 1 alone, and only 
`r am_ecology_percent`% of the spawning stock biomass that would have been 
achieved under Model 1 alone.

To plot growth curves of individual models (Fig 4a), we evaluate

$$\Delta x = x_{t+1} - x_t = f(x_t) - x_t$$

for each model for all possible states $x_t$.

```{r plot_models}
model_curves <- 
  map_dfc(models, function(f) f(states) - states) %>%
  mutate(state = states) %>%  pivot_longer(names(models), "model")

write_csv(model_curves, "../data/model_curves.csv")

@@

### 2021-03-01 manuscript/manuscript.Rmd
data cache

---
title: 'Learning to be wrong:'
author:
- name: Carl Boettiger
  email: cboettig@berkeley.edu
  affiliation: a,1
date: "`r Sys.Date()`"

@@

# For footer text
lead_author_surname: Boettiger

significance: |

@@

output: rticles::pnas_article
---

@@

I illustrate this problem using an example from fisheries management. 
Fisheries are a significant economic and conservation concern world
wide and their management remains an important debate [e.g. @Worm2006;
@Worm2009; @Costello2016]. Moreover, their management has been a proving 
grounds for theoretical and practical decision-making issues which are widely
applicable in other areas of ecology and conservation [@Ludwig1993; @Lande1994],
and one that has long wrestled with issues of uncertainty in the context
of management decisions [e.g. @Clark1973; @Reed1979; @Walters1981; @Ludwig1982].
While modern fisheries management frequently relies on complex models which may
contain scores of parameters to reflect the specific age or stage structure of a
specific fish stock [@ram], I will rely on simple, well-studied models which
permit greater intuition and generalization [@Levins1966]. Consistent with 
such previous work [@Schaefer1954; @Clark1973; @Reed1979; @Walters1981; @Ludwig1982; @Costello2016],
let us consider the problem of determining the optimal harvest policy given a 
measurement of the current stock size.

@@

Further we imagine that the function $f$ is not known precisely, and so we will rely on an evaluation of forecasting skill across a set of candidate models to determine which one to use to manage the fishery.  Again for simplicity, we will restrict ourselves to two simple candidate models $f_1$ and $f_2$.  Both share the same underlying structure of logistic recruitment (known as the Gordon-Schaefer model in fisheries context owing to groundbreaking work independently by @Schafer1954 and @Gordon1954), differing only in their choice of certain parameters:

@@

A wide range of paradigms are available for approaching the issue of decision-making under uncertainty. These approaches can roughly be divided into two groups: the first group treats the issue of model uncertainty independently from the decision itself, while the second integrates the process of reducing model uncertainty into the process of decision making to maximize the value of some objective. There are a wide range of techniques within each, and it is also possible to blend approaches. The key distinction is that methods in the first group do not involve any direct consideration of the possible actions or the utility that may result from those actions in how they select models [statistical approaches such as information criteria @BurnhamAnderson; and in particular, forecasting evaluation @Clark2001; @Dietze2018], while those in the second group require a more explicit statement of possible actions and the desired objectives [Decision theoretic approaches, for which @Polasky2011 provides an excellent and accessible review]. I illustrate how the most promising techniques from each of these groups would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcomes.  In retrospect, it will become obvious that neither these nor any other widely applied methods will select the model that leads to the best outcomes from the set of models considered.

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios. In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under Model 1 and Model 2 respectively [Fig].  The mean, $\mu$ and variance, $\sigma$ of the forecast are compared against the true observation $x$ using a Proper scoring rule of @Geiting2007,

\bein{equation}
-(\mu - x )^2 / \sigma^2  - \log(\sigma) \label{proper}
\end{equation}

for each prediction over 100 replicate simulations of 100 time steps each [Fig].

```{r figure1, fig.width=7, fig.height=5, fig.cap = "Forecast performance of each model.  Panels A, B: Step ahead predictions of stock size under unfished (A) and fished (B) scenarios. Error bars indicating the 95% confidence intervals around each prediction, while stars denote the observed value in that year.  Because the models make different decisions each year in the fished scenario, the observed stock size in year 2, 3, etc under the management of model 1 (blue stars) is different from that under model 2 (red stars).  Panels C, D: corresponding distribution of proper scores across all predictions (100 replicates of 100 timesteps). Higher scores are better, confirming that model 2 makes the better forecasts."}
fig1ab + fig1cd + 
  plot_layout(widths =  c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

In Scenario B (actively harvest), I have first solved for the optimal management strategy for both Model 1 and Model 2 [Fig], which for small noise and concave functions with linear reward structure can be done analytically [see proof in @Reed1979], or can be solved more generally by stochastic dynamic programming [see review by @Marescot2013, details in Appendix].  Under this scenario, replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either Model 1 and Model 2. The resulting stock sizes in the time-step following this harvest are once again compared to the probabilities predicted by each model using Eq \eqref{proper}. Model 2 unequivocally outperforms Model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of Model 2 in both scenarios, the outcomes from management under Model 2 are substantially worse.  We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass). In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.  This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests. While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, Model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig ].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from Model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under Model 1 are also substantially higher (panel B)"}
fig2a + fig2b
```

@@

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the simulation. Research has long emphasized the importance of learning and adaption in the face of new data whenever we are dealing with model uncertainty [@Polasky2011]. A related limitation is that in Scenario B, decisions were based either on assuming Model 1 is correct or assuming Model 2 is correct. A more robust approach [e.g. @Walters1978] incorporates the uncertainty over models directly into the decision calculations by integrating over the probability each model $M$ (or parameter value) is correct when calculating the utility; $U(x_t, a_t) = \int U(x_t, a_t | M) P(M)$. The probability assigned to each model can then be updated after each subsequent action in light of the resulting outcomes [@Ludgwig1982, @Smith1981]. While a the wide range of strategies for such iterative updating are known [@Polasky2011]; all will fail to select the higher-performing of the two simple models considered  here.

@@

Each of these approaches swiftly decides that model 1 in our example is implausible, and settles for making decisions driven almost exclusively by model 2.  I illustrate this using a passive adaptive management strategy following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty. To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating [Fig , achieving a net present value over 100 time steps that only 31% that expected under management using Model 1 alone]. If instead of considering only these two models we consider a whole suite of Gordon-Schaefer-type models with varying $r$ and $K$ parameters, learning is slightly slower but no less counter-productive: mean stock size is $46%$ and net present value is $20%$ of what would be achieved under Model 1 (Appendix, Fig S1).

```{r}
am <- am1$df %>% mutate(belief = am1$posterior$V2)
am %>%  ggplot(aes(time, states[state], col = belief)) + 
  geom_line() + 
  geom_point() +
  scale_colour_gradient(limits = c(0,1), low = pal[1], high = pal[4]) +
  ylab("stock size")
```

@@

```{r fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth."}
# Fig 4
plot_models + plot_policies
```

>>>>

---
title: 'Are we learning to build better models at the cost of better decisions?'
author:
- name: Carl Boettiger
  email: cboettig@berkeley.edu
  affiliation: a,1
date: "`r Sys.Date()`"

@@

@@

output: rticles::elsevier_article
layout: 3p
---

@@

I illustrate this problem using an example from fisheries management. 
Fisheries are a significant economic and conservation concern world
wide and their management remains an important debate [e.g. @Worm2006;
@Worm2009; @Costello2016]. Moreover, their management has been a proving 
grounds for theoretical and practical decision-making issues which are widely
applicable in other areas of ecology and conservation [@Ludwig1993; @Lande1994],
and one that has long wrestled with issues of uncertainty in the context
of management decisions [e.g. @Clark1973; @Reed1979; @Walters1981; @Ludwig1982].
While modern fisheries management frequently relies on complex models which may
contain scores of parameters to reflect the specific age or stage structure of a
specific fish stock [@ramlegacy2018; @ram], I will rely on simple, well-studied models which
permit greater intuition and generalization [@Levins1966]. Consistent with 
such previous work [@Schaefer1954; @Clark1973; @Reed1979; @Walters1981; @Ludwig1982; @Costello2016],
let us consider the problem of determining the optimal harvest policy given a 
measurement of the current stock size.

@@

Further we imagine that the function $f$ is not known precisely, and so we will rely on an evaluation of forecasting skill across a set of candidate models to determine which one to use to manage the fishery.  Again for simplicity, we will restrict ourselves to two simple candidate models $f_1$ and $f_2$.  Both share the same underlying structure of logistic recruitment (known as the Gordon-Schaefer model in fisheries context owing to groundbreaking work independently by @Schaefer1954 and @Gordon1954), differing only in their choice of certain parameters:

@@

A wide range of paradigms are available for approaching the issue of decision-making under uncertainty. These approaches can roughly be divided into two groups: the first group treats the issue of model uncertainty independently from the decision itself, while the second integrates the process of reducing model uncertainty into the process of decision making to maximize the value of some objective. There are a wide range of techniques within each, and it is also possible to blend approaches. The key distinction is that methods in the first group do not involve any direct consideration of the possible actions or the utility that may result from those actions in how they select models [statistical approaches such as information criteria e.g. @Burnham1998]; and in particular, forecasting evaluation @Clark2001; @Dietze2018], while those in the second group require a more explicit statement of possible actions and the desired objectives [Decision theoretic approaches, for which @Polasky2011 provides an excellent and accessible review]. I illustrate how the most promising techniques from each of these groups would be applied to this simple problem, and demonstrate that it both cases they lead us away from the model that produces the most desirable decisions towards worse outcomes.  In retrospect, it will become obvious that neither these nor any other widely applied methods will select the model that leads to the best outcomes from the set of models considered.

@@

I illustrate the process of model selection by strictly proper scoring rules using two scenarios. In Scenario A (passive observation) the fish stock is unharvested and allowed to recover towards carrying capacity (as simulated under our "true" model, Model 3) while comparing the observed stock size in each subsequent time step to the distribution predicted under Model 1 and Model 2 respectively [Fig].  The mean, $\mu$ and variance, $\sigma$ of the forecast are compared against the true observation $x$ using a Proper scoring rule of @Gneiting2007,

\begin{equation}
-(\mu - x )^2 / \sigma^2  - \log(\sigma) \label{proper}
\end{equation}

for each prediction over 100 replicate simulations of 100 time steps each [Fig 1].

```{r figure1, fig.width=7, fig.height=5, fig.cap = "Forecast performance of each model.  Panels A, B: Step ahead predictions of stock size under unfished (A) and fished (B) scenarios. Error bars indicating the 95% confidence intervals around each prediction, while stars denote the observed value in that year.  Because the models make different decisions each year in the fished scenario, the observed stock size in year 2, 3, etc under the management of model 1 (blue stars) is different from that under model 2 (red stars).  Panels C, D: corresponding distribution of proper scores across all predictions (100 replicates of 100 timesteps). Higher scores are better, confirming that model 2 makes the better forecasts."}
predictions <- read_csv("../data/predictions.csv") %>% mutate(model = as.character(model))

fig1cd <- predictions %>%
  ggplot(aes(x = score, group = model, fill = model)) +
  geom_histogram(binwidth = 2, show.legend = FALSE) +
  coord_cartesian(xlim = c(-100, 1), ylim = c(0,4000)) +
  xlab("Proper score") +
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "C.",
    B_fished = "D."))) + 
  scale_x_continuous(breaks = c(-100, 0)) +
  theme(axis.text.y = element_blank(),
        plot.margin = margin(0, 0, 0, 0, "cm"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()
        )

 fig1ab <- predictions %>% 
  filter(reps == "2", time < 10) %>%
  ggplot(aes(time, col = model, fill = model)) + 
  geom_point(aes(y = expected)) + 
  geom_errorbar(aes(ymin = low, ymax = high)) +
  geom_point(aes(y = true), pch = "*", size = 12, 
             alpha = 0.6) + 
  facet_wrap(~scenario, ncol = 1, labeller =as_labeller(c(
    A_unfished = "A. Without fishing",
    B_fished = "B. Managed harvest"))) +
  ylab("stock size")

 fig1ab + fig1cd + 
  plot_layout(widths =  c(3,1), guides = 'collect') + 
  theme(plot.margin = margin(0, 0, .1, .1, "cm"))
```

In Scenario B (actively harvest), I have first solved for the optimal management strategy for both Model 1 and Model 2 [Fig 1b]. For small noise and concave functions with linear reward structure this can be done analytically [see proof in @Reed1979], or solved more generally by stochastic dynamic programming [see review by @Marescot2013, details in Appendix].  Under this scenario, replicate simulations of the stock are harvested at each time step using the optimal quota dictated by either Model 1 and Model 2. The resulting stock sizes in the time-step following this harvest are once again compared to the probabilities predicted by each model using Eq \eqref{proper}. Model 2 unequivocally outperforms Model 1 in both scenarios of passive observation and active harvest.

Despite the clearly superior predictive accuracy of Model 2 in both scenarios, the outcomes from management under Model 2 are substantially worse.  We can assess such outcomes in less abstract terms than forecasting skill, such as economic value (in dollars) or the ecological value (unharvested biomass). In our simple formulation of the decision problem, the "utility" the manager seeks to maximize is simply the economic value (net present value: the discounted sum of all profits from future harvests, Eq \eqref{utility}) of harvested fish.  This formulation ignores any utility provided by fish that are not harvested, beyond their contribution to future potential harvests. While it is possible to include such contributions directly in the utility function being optimized [e.g. @Halpern2013], even without doing so, Model 1 maintains both a higher unharvested biomass and also leads to higher economic returns throughout [Fig 2].

```{r figure2, message = FALSE, fig.cap = "Ecological and economic performance of each forecast. Harvest quotas derived from Model 1 result in a significantly higher fish stock size than under Model 2 (panel A). Economic returns under Model 1 are also substantially higher (panel B)"}

sims <- read_csv("../data/sims.csv")
npv_df <- read_csv("../data/npv_df.csv")

plot_ecology <- 
  sims %>%
  mutate(model = as.character(model), resp = as.character(reps)) %>%
  group_by(model, time) %>%
  summarise(mean_state = mean(state), sd = sd(state), .groups = "drop") %>%
  filter(time < 25, model != "3") %>%
  ggplot(aes(time, mean_state)) + 
  geom_line(aes(col = model), lwd=1.5, show.legend = FALSE) +
    geom_ribbon(aes(ymin = mean_state - 2*sd, 
                    ymax = mean_state + 2*sd,
                    fill = model),
                alpha = 0.2, show.legend = FALSE) +
  ylab("state") + 
  labs(subtitle = "A. Ecological outcomes")

optimal <- select(filter(npv_df, model == "3"), time, mean_npv)

plot_econ <- 
npv_df %>%
  mutate(model = as.character(model), resp = as.character(reps)) %>%
  filter(model != "3", time %in% seq(1,100, by = 5)) %>%
  ggplot(aes(time, mean_npv)) +
  geom_line(data = optimal, lwd = 1.5, col = "grey20") +
  geom_point(aes(col=model), size = 4, alpha = 0.8) + 
  ylab("Net present value") + xlab("time")  + 
  labs(subtitle = "B. Economic outcomes")

plot_ecology + plot_econ
```

@@

One obvious limitation in this comparison is that scenario B treats each model as fixed over the entire course of the simulation. Research has long emphasized the importance of learning and adaption in the face of new data whenever we are dealing with model uncertainty [@Polasky2011]. A related limitation is that in Scenario B, decisions were based either on assuming Model 1 is correct or assuming Model 2 is correct. A more robust approach [e.g. @Walters1978] incorporates the uncertainty over models directly into the decision calculations by integrating over the probability each model $M$ (or parameter value) is correct when calculating the utility; $U(x_t, a_t) = \int U(x_t, a_t | M) P(M)$. The probability assigned to each model can then be updated after each subsequent action in light of the resulting outcomes [@Ludwig1982, @Smith1981]. While a the wide range of strategies for such iterative updating are known [@Polasky2011]; all will fail to select the higher-performing of the two simple models considered  here.

@@

Each of these approaches swiftly decides that model 1 in our example is implausible, and settles for making decisions driven almost exclusively by model 2.  I illustrate this using a passive adaptive management strategy following classic examples for parameter [@Ludwig1982] or structural [@Smith1981] model uncertainty. To demonstrate that the behavior is not driven by failure to explore sufficiently, (which might be addressed by an active adaptive management), I will assign initial probability that model 2 is true at 1%.  After a single iteration of learning, these probabilities are completely reversed, with the manager deciding that model 2 is almost certainly correct  [Fig]. As before, this results in a management practice with much worse ecological and economic outcomes than would have been realized by a manager who stubbornly clung to model 1 without updating [Fig 2, achieving a net present value over 100 time steps that only 31% that expected under management using Model 1 alone]. If instead of considering only these two models we consider a whole suite of Gordon-Schaefer-type models with varying $r$ and $K$ parameters, learning is slightly slower but no less counter-productive: mean stock size is $46%$ and net present value is $20%$ of what would be achieved under Model 1 (Appendix, Fig S1).

```{r figure3, fig.cap="Adaptive management under model uncertainty, color indicates the belief that Model 2 is correct (red). Each model is assigned an initial prior belief. To ensure that any issues with passive exploration are not the cause of preferring Model 2, the initial belief in Model 2 is set to 1%.  Within a single iteration of adaptive management, the belief over models is updated to near certainty in Model 2, resulting in higher harvests and lower stock sizes similar to managing under Model 2 alone, Fig 2."}
am <- read_csv("../data/am.csv")
am %>%  filter(time <30) %>%
  ggplot(aes(time, states[state], col = belief)) + 
  geom_line() + 
  geom_point() +
  scale_colour_gradient(limits = c(0,1), low = pal[1], high = pal[4]) +
  ylab("stock size")
```

@@

```{r figure4,  fig.cap="Panel A: Population growth curves of each model. The positive equilibrium of each model occurs where the curve crosses the horizontal axis. Note that while Model 2 is a better approximation to the truth (Model 3), Model 1 better approximates the stock size which leads to maximum growth.", fig.width=8}
# Fig 4
model_curves <- read_csv("../data/model_curves.csv") %>% mutate(model = as.character(model))
policies <- read_csv("../data/policies.csv") %>% mutate(model = as.character(model))

plot_models <- 
  model_curves %>% 
  ggplot(aes(state, value, col=model, lty=model)) +
  geom_hline(aes(yintercept = 0), lwd=1) + 
  geom_line(lwd=2, show.legend = FALSE) + 
  coord_cartesian(ylim = c(-5, 8), xlim = c(0,16)) +
  ylab(bquote(f(x) - x)) + xlab("x")

plot_policies <- policies %>%
  ggplot(aes(states, escapement, col=model, lty=model)) + 
  geom_line(lwd=2) + xlab("state")

plot_models + labs(subtitle="A") +  plot_policies +  labs(subtitle="B")
```
